### 2022-10-04 PSTAT131_Final_Project.Rmd
progress

**Loading the dataset:**  
```{r}
pokemon <- read.csv("pokemon.csv")
head(pokemon)
```

Outcome-  
capture_rate: capture rate

https://bulbapedia.bulbagarden.net/wiki/Stat
Variables of interest-  
attack (how much damage they deal using a physical move)  
classification 
defense 
hp (hit points - how much damage they can receive before fainting)
sp_attack (how much damage the deal using special moves)   
sp_defense  
type1  
type2  
generation
is_legendary

```{r}
library(tidyverse)
ggplot(pokemon, aes(x=attack, y=capture_rate)) + geom_point()
```

```{r}

library(dplyr)
filter(pokemon, capture_rate %in% c("30 (Meteorite)255 (Core)"))
pokemon[pokemon$name=="Minior", "capture_rate"] <- 30
# I chose 30 because the capture rate only rises to 255 if you break its shell
filter(pokemon, name %in% c("Minior"))
```

```{r}
ggplot(pokemon, aes(x=attack, y=capture_rate)) + geom_point()
pokemon[, "capture_rate"] <- as.numeric(pokemon$capture_rate)
```

```{r}
ggplot(pokemon, aes(x=attack, y=capture_rate)) + geom_point()
ggplot(pokemon, aes(x=hp, y=capture_rate)) + geom_point()
ggplot(pokemon, aes(x=generation, y=capture_rate, group=generation)) + geom_boxplot()
ggplot(pokemon, aes(x=defense, y=capture_rate)) + geom_point()
ggplot(pokemon, aes(x=sp_attack, y=capture_rate)) + geom_point()
ggplot(pokemon, aes(x=sp_defense, y=capture_rate)) + geom_point()
ggplot(pokemon, aes(x=classfication, y=capture_rate, group=classfication)) + geom_boxplot()
ggplot(pokemon, aes(x=type1, y=capture_rate, group=type2)) + geom_boxplot()
```