### 2022-10-04 PSTAT131_Final_Project.Rmd
edit

**Loading the dataset:**  
```{r}
pokemon <- read.csv("pokemon.csv")
head(pokemon)
```

Outcome-  
capture_rate: capture rate

https://bulbapedia.bulbagarden.net/wiki/Stat
Variables of interest-  
attack (how much damage they deal using a physical move)  
classification 
defense 
hp (hit points - how much damage they can receive before fainting)
sp_attack (how much damage the deal using special moves)   
sp_defense  
type1  
type2  
generation
is_legendary

```{r}
library(tidyverse)
ggplot(pokemon, aes(x=attack, y=capture_rate)) + geom_point()
```

```{r}

library(dplyr)
filter(pokemon, capture_rate %in% c("30 (Meteorite)255 (Core)"))
pokemon[pokemon$name=="Minior", "capture_rate"] <- 30
# I chose 30 because the capture rate only rises to 255 if you break its shell
filter(pokemon, name %in% c("Minior"))
```

```{r}
ggplot(pokemon, aes(x=attack, y=capture_rate)) + geom_point()
pokemon[, "capture_rate"] <- as.numeric(pokemon$capture_rate)
```

```{r}
ggplot(pokemon, aes(x=attack, y=capture_rate)) + geom_point()
ggplot(pokemon, aes(x=hp, y=capture_rate)) + geom_point()
ggplot(pokemon, aes(x=generation, y=capture_rate, group=generation)) + geom_boxplot()
ggplot(pokemon, aes(x=defense, y=capture_rate)) + geom_point()
ggplot(pokemon, aes(x=sp_attack, y=capture_rate)) + geom_point()
ggplot(pokemon, aes(x=sp_defense, y=capture_rate)) + geom_point()
ggplot(pokemon, aes(x=classfication, y=capture_rate, group=classfication)) + geom_boxplot()
ggplot(pokemon, aes(x=type1, y=capture_rate, group=type2)) + geom_boxplot()
```

```{r}
pokemon %>%
  distinct(classfication)
```

Might not be a good idea to do classification because there are 588 unique classifications of a pokemon --> better to focus on type

```{r}
library(corrplot)
pokemon_numeric <- pokemon %>%
  select_if(is.numeric)
# taking out the aforementioned columns with too many missing values (was not planning on using them anyways)
pokemon_numeric <- pokemon_numeric[, !names(pokemon_numeric) %in% c("height_m", "weight_kg", "percentage_male")]

pokemon_cor <- cor(pokemon_numeric)
pokemon_corrplot <- corrplot(pokemon_cor)
```

We can see that how a pokemon performs against a certain other type of pokemon is not really correlated to capture rate much. However, this also shows how capture rate has a high negative correlation with base_total which is the Base Stat Total (BST = hp + attack + defense + sp_attack + sp_defense + speed). So, it might be worth exploring how all of those added together affects the capture rate.

```{r}
ggplot(pokemon, aes(capture_rate)) + geom_histogram()
```

No specific skewness is visible, but there is a peak in the data around 40-50 capture rate.