### 2022-10-08 Draft.Rmd
loading and exploring

## Loading and Exploring Raw Data

```{r}
# Loading all necessary packages
library(tidyverse)
library(dplyr)
library(tidymodels)
```

Now, let's load our data and start our exploration to see what we're working with!

```{r}
# Assigning the data to a variable
pokemon <- read.csv("data/pokemon.csv")

# Calling head() to see the first few rows
head(pokemon)
```

How many rows and columns are we working with?

```{r}
pokemon %>%
  dim()
```

We have a total of 801 rows, one for each Pokemon in the dataset, and 41 columns, which are the attributes reported about each Pokemon. Within these attributes, we have our response variable, `capture_rate`, as well as 40 features. Obviously, we will not be working with all 40 features so we will need to continue exploring the variables to decide which ones should be kept and which should not.

From the data, each of the variables that specify each Pokemon's skill against a certain type of Pokemon (i.e. `against_bug`, `against_dark`, etc.) do not appear to be as relevant as other attributes in deciding whether specifc features about a Pokemon affect its capture rate. We can also assume that the `attack` variable gives us an overall estimate of each Pokemon's attack skill so we would not need to go into such great detail as finding out how the Pokemon perform against every other type of Pokemon. For that reason, these 18 columns can be dropped.

At first, it seems that the `classification` variable may be helpful; however, let us see how many unique classifications of a Pokemon there are.

```{r}
pokemon %>%
  distinct(classfication) %>%
  count()
```

There are 588 unique classifications! That seems a little too high for our analysis so we will just stick to using the `type` of the Pokemon instead.

Furthermore, for the `abilities` variable, we can see that there is no set amount for how many abilities a Pokemon has or a measure to figure out which abilities are more powerful than others. For these reasons, it may be best to avoid including this column into our analysis.

Also, the `base_happiness`, `base_egg_steps`, and `experience_growth` variables appear to be the same across most Pokemon as they are base statistics that each Pokemon shares. These variables would not be helpful in our analysis since we are trying to find how the major differences between each Pokemon affect their capture rate.

Some other variables must also be taken out as they do not seem as if they would reveal any more information relevant to our analysis: `weight_kg`, `pokedex_number`, `percentage_male`, `japanese_name`, and `height_m`.

## Morphing and Tidying Data