### 2022-11-07 Draft2.Rmd
progress

```{r}
# Loading all necessary packages
library(tidyverse)
library(dplyr)
library(tidymodels)
library(readr)
# Assigning the data to a variable
pokemon <- read_csv("data/pokemon.csv")
tidymodels_prefer()
set.seed(0124)
```

```{r}
# Calling head() to see the first few rows
head(pokemon)
```

```{r}
# Calling dim() to see how many rows and columns
dim(pokemon)
```

```{r}
# Selecting only the variables we want
pokemon <- pokemon %>% 
  select(c("abilities", "attack", "base_total", "capture_rate", "defense", "hp", "name", "sp_attack", "sp_defense", "speed", "type1", "generation", "is_legendary"))
head(pokemon)
```

```{r}
# summary stats
pokemon %>% 
  summary()

# missing values?
colSums(is.na(pokemon))
```

```{r}
# capture rate should not be a character
library(dplyr)
filter(pokemon, capture_rate %in% c("30 (Meteorite)255 (Core)"))
pokemon[pokemon$name=="Minior", "capture_rate"] <- "30"
pokemon$capture_rate <- as.integer(pokemon$capture_rate)
# I chose 30 because the capture rate only rises to 255 if you break its shell
filter(pokemon, name %in% c("Minior"))
```

```{r}
generation_levels <- c(1,2,3,4,5,6,7)
pokemon$generation <- factor(pokemon$generation, levels=generation_levels)
# MAKE IS_LEGENDARY A FACTOR
```

```{r}
pokemon %>% 
  summary()
```

```{r}
library(tidytext)
```

```{r}
test <- pokemon$abilities %>% 
  str_split(pattern = "', '") #%>% 
  # str_extract_all(string = ., pattern = "\\w+", simplify = TRUE)
  # map(str_extract_all(string = ., pattern = "\\w+", simplify = TRUE))
all_abilities <- test %>% map(~ str_replace_all(string = ., pattern = "[[:punct:]]", "")) %>% 
  unlist() %>% as.factor() %>% 
  data.frame(abilities = .)

all_abilities %>% group_by(abilities) %>% summarise(count = n(), capture_rate = mean(capture_rate)) %>% 
  arrange(desc(count)) %>% View()

#ggplot(data = tibble(all_abilities)) +
 # geom_bar(aes(x = all_abilities))
```

```{r}
pokemon_copy <- pokemon
pokemon_copy$abilities <- pokemon_copy$abilities %>% 
  str_split(pattern = "', '") %>% 
  map(~ str_replace_all(string = ., pattern = "[[:punct:]]", ""))

#tidy_abilities <- pokemon_copy %>% 
#  unnest_tokens(words, abilities)

pokemon_copy2 <- pokemon_copy %>% 
  unnest()

pokemon_copy2 %>% group_by(abilities) %>%  summarise(count = n(), capture_rate = mean(capture_rate)) %>% 
  arrange(desc(count)) %>% View()
```

```{r}
library(dplyr)

pokemon_abilities_grouped <- pokemon_copy2 %>% 
  group_by(abilities)

pokemon_abilities_grouped %>% 
  summarise(
    capture_rate = mean(capture_rate)
  ) %>% 
  arrange(desc(capture_rate))

```

```{r}
mean(pokemon_copy2$capture_rate)
```

```{r}
# making boolean columns for if a pokemon has the top 5 most apparent abilities
pokemon_copy3 <- pokemon

sturdy <- "Sturdy"
pokemon_copy4 <- pokemon %>% 
  mutate(has_sturdy = if_else(grepl("Sturdy", pokemon$abilities), 1, 0)) %>% 
  mutate(has_swift_swim = if_else(grepl("Swift Swim", pokemon$abilities), 1, 0)) %>% 
  mutate(has_keen_eye = if_else(grepl("Keen Eye", pokemon$abilities), 1, 0)) %>% 
  mutate(has_chlorophyll = if_else(grepl("Chlorophyll", pokemon$abilities), 1, 0)) %>% 
  mutate(has_levitate = if_else(grepl("Levitate", pokemon$abilities), 1, 0))

sum(pokemon_copy4$has_sturdy)
sum(pokemon_copy4$has_swift_swim)
sum(pokemon_copy4$has_keen_eye)
sum(pokemon_copy4$has_chlorophyll)
sum(pokemon_copy4$has_levitate)

write_csv(pokemon_copy4, "pokemon_final.csv")
```

```{r}
pokemon_data <- read_csv("pokemon_final.csv")
head(pokemon_data)
```

```{r}
# Changing categorical predictors to factors
pokemon_data$type1 <- as.factor(pokemon_data$type1)
pokemon_data$generation <- as.factor(pokemon_data$generation)
pokemon_data$is_legendary <- as.factor(pokemon_data$is_legendary)
pokemon_data$has_sturdy <- as.factor(pokemon_data$has_sturdy)
pokemon_data$has_swift_swim <- as.factor(pokemon_data$has_swift_swim)
pokemon_data$has_keen_eye <- as.factor(pokemon_data$has_keen_eye)
pokemon_data$has_chlorophyll <- as.factor(pokemon_data$has_chlorophyll)
pokemon_data$has_levitate <- as.factor(pokemon_data$has_levitate)
```

```{r}
summary(pokemon_data)
```

```{r}
# Splitting the data
pokemon_split <- initial_split(pokemon_data, prop = 0.7, strata = capture_rate)
pokemon_train <- training(pokemon_split)
pokemon_test <- testing(pokemon_split)
```

```{r}
# Creating recipe
pokemon_recipe <- recipe(capture_rate ~ attack + defense + hp + sp_attack + sp_defense + speed + type1 + generation + is_legendary + has_sturdy + has_swift_swim + has_keen_eye + has_chlorophyll + has_levitate, data = pokemon_train) %>% 
  step_dummy(all_nominal_predictors()) %>% 
  step_center(all_predictors()) %>% 
  step_scale(all_predictors())
```

```{r}
# Creating folds
pokemon_folds <- vfold_cv(pokemon_train, v = 10)
```

```{r}
# Linear regression 
lm_model <- linear_reg() %>% 
  set_engine("lm")

lm_workflow <- workflow() %>% 
  add_model(lm_model) %>% 
  add_recipe(pokemon_recipe)
```

```{r}
# Fitting the models
lm_fit <- fit_resamples(lm_workflow, resamples = pokemon_folds)
```
```{r}
collect_metrics(lm_fit)
```